---
title: "Data Details"
---

## Variable Descriptions

| Variable Name      | Description                                                                                                                                      |
|--------------------|--------------------------------------------------------------------------------------------------------------------------------------------------|
| `release`          | The release. Current version is V0.1 (Preprint).                                                                                                 |
| `release_date`     | The release date. Latest version is 2023-12-18. This is updated manually with changes to the data or model that create changes in the estimates. |
| `data_source`      | The source of the events data. We have IDEA and ICEWS versions.                                                                                  |
| `created_date`     | The date that the dissent scores were created. This updates with every fit of the model.                                                         |
| `country_name`     | The name of the country.                                                                                                                         |
| `ccode`            | The numeric COW code for the country.                                                                                                            |
| `stateabb`         | The 3-digit alpha COW code for the country.                                                                                                      |
| `year`             | The year.                                                                                                                                        |
| `n_events`         | The total number of events (dissent and not-dissent).                                                                                            |
| `n_dissent_events` | The total number of *dissent* events in the country-year.                                                                                        |
| `avg_pi`           | The posterior average of the parameter $\pi$ from the statistical model.                                                                         |
| `avg_eta`          | The posterior average of the parameter $\eta$ from the statistical model.                                                                        |
| `dissent_score`    | The dissent score. This is $\eta$ rescaled to have mean = 0 and SD = 0.5.                                                                        |
| `se_dissent_score` | The posterior SD of the dissent score.                                                                                                           |

## Summary Using `glimpse()`

```{r}
#| message: false

# load packages
library(tidyverse)
library(dataverse)

# get dissent scores from dataverse
dissent <- get_dataframe_by_name(
  filename = "dissent-scores.tab",
  dataset  = "doi:10.7910/DVN/CL4CA8",
  server   = "dataverse.harvard.edu", 
  original = TRUE, 
  .f = readr::read_csv)

# quick look
glimpse(dissent)
```

## Summary Using `skim()`

```{r}
# generate numerical summaries
skimr::skim(dissent)
```
