---
title: "Data Details"
---

## Variable Descriptions

| Variable Name      | Description                                                                      |
|--------------------|----------------------------------------------------------------------------------|
| `release`          | The release. Current version is V0.0.9999, an alpha release for initial testing. |
| `release_date`     | The release date. Latest version is 2023-09-13.                                  |
| `country_name`     | The name of the country.                                                         |
| `ccode`            | The numeric COW code for the country.                                            |
| `stateabb`         | The 3-digit alpha COW code for the country.                                      |
| `year`             | The year.                                                                        |
| `n_events`         | The total number of events (dissent and not-dissent).                            |
| `n_dissent_events` | The total number of *dissent* events in the country-year.                        |
| `avg_pi`           | The posterior average of the parameter $\pi$ from the statistical model.         |
| `avg_eta`          | The posterior average of the parameter $\eta$ from the statistical model.        |
| `dissent_score`    | The dissent score. This is $\eta$ rescaled to have mean = 0 and SD = 0.5.        |
| `se_dissent_score` | The posterior SD of the dissent score.                                           |

## Summary Using `glimpse()`

```{r}
#| message: false

# load packages
library(tidyverse)

# read csv directly from github
gh_url <- "output/dissent-scores.csv"
dissent <- read_csv(gh_url)

# quick look
glimpse(dissent)
```

## Summary Using `skim()`

```{r}
# generate numerical summaries
skimr::skim(dissent)
```
